### 2015-02-10 Studies/2012_PG_County/Report/PG_Co_Fire.tex
PG Co: update report, plotting script, and figure | Bib: update reference

\large
\flushright{\bf National Institute of Standards and Technology Technical Note XXXX \\
Natl.~Inst.~Stand.~Technol.~Tech.~Note~XXXX, \pageref{LastPage} pages (November 2014) \\
CODEN: NTNOEF}

@@

\begin{table}[!h]
\centering
\captionof{table}[Abridged approximate fire event timeline]{Abridged Prince George's Fire/EMS approximate fire event timeline~\cite{PGCounty2013}.}\label{tab:fire_info}
\begin{tabular}{cl}
\toprule[1.5pt]
Incident Time              &  Fire Behavior / Fireground Operation  \\
{[hh:mm:ss]}               &                                        \\
\midrule
21:08:26   & First 911 call is made. \\[.25cm]
\multirow{3}{*}{21:11:03}  &   \multirow{3}{*}{\parbox{10cm} {Dispatch for a ``box'' alarm assignment- four engine companies, two truck companies, one rescue squad, and a commanding officer.}}    \\
                           &         \\
                           &         \\[.25cm]
21:12:55  & First unit (Engine 807B) arrived on-scene.  \\[.25cm]
21:13:44  & Volunteer Chief 809A establishes incident command (IC).   \\[.25cm]
21:13:56  & Truck 809 arrived on-scene. \\[.25cm]
21:14:00  & Front door opened. \\[.25cm]
21:14:42  & Truck 809 personnel enter structure. \\[.25cm]
21:15:47  & Basement door opened and water applied to fire. Front door blown closed. \\[.25cm]
21:15:51  & Lower pane of front egress window broken. \\[.25cm]
21:16:01  & Top pane of front egress window broken. \\[.25cm]
21:16:16  & Truck 809 officer bailout through front window. \\[.25cm]
21:16:44  & Front door opened. \\[.25cm]
21:17:15  & Truck 809 forcible entry firefighter rescued. \\[.25cm]
\bottomrule[1.25pt]
\end{tabular}\par
\end{table}

@@

\begin{figure}[!ht]
\includegraphics[width=0.75\textwidth]{../Figures/Fire_HRR}
\caption[Prescribed HRR vs. time for the simulation.]
{Prescribed HRR vs. time for the simulation. The text labels indicate the prescribed steady peak HRR and the decay phase associated with firefighter water application.}
\label{fig:hrr}
\end{figure}

@@

In addition to the state of openings within the structure, outlined in Table~\ref{tab:vents}, an 8.9~m/s (20~mph) wind velocity boundary condition was assumed. The flow direction of the wind was from the rear toward the front of the structure, perpendicular to the rear of the structure.  Based on local weather conditions, mean wind velocities ranged between 4.5~m/s (10~mph) and 8.9~m/s (20~mph) with gusts up to 13.4~m/s (30~mph) with a direction that angled toward the back left of the structure~\cite{PGCounty2013}.

\begin{table}
\centering
\captionof{table}{Timeline of ventilation changes in the simulation}\label{tab:vents}
\begin{tabular}{lllll}
\toprule[1.5pt]
Baseline     &   Ventilation   &  Type       & State of       &  Side of  \\
Simulation  &  Area             &                 & Opening     &  Structure  \\
{[s]}             &  {[m$^2$]}     &                 &           &               \\
\midrule
0               &  1.40             &  Window      &  Opened     &  Rear  \\[.25cm]
0               &  0.39             &  Window      &  Opened     &  Left  \\[.25cm]
100             &  1.84             &  Door        &  Opened     &  Front \\[.25cm]
207             &  1.84             &  Door        &  Closed     &  Front \\[.25cm]
207             &  1.84             &  Door        &  Opened     &  Rear  \\[.25cm]
211             &  0.26             &  Window      &  Opened     &  Front \\[.25cm]
221             &  0.26             &  Window      &  Opened     &  Front \\[.25cm]
267             &  1.84             &  Door        &  Opened     &  Front \\[.25cm]

@@

\begin{table}[!ht]
\centering
\captionof{table}{Relevant fire model input parameters.}\label{tab:mod_param}
\begin{tabular}{lll}
\toprule[1.5pt]
Parameter                                                  & Description                                   & Discussion                            \\
\midrule
Simulation Time                                            &  5~min                                        & --                                    \\ [.25cm]
\multirow{2}{*}{Grid Cell Size}                            &  10~cm around house                           & \multirow{2}{*}{Section~\ref{mesh}}   \\ 
                                                           &  20~cm for wind                               &                                       \\ [.25cm]
Ambient Temperature$^*$                                    &  20~$^{\circ}$C (68~$^{\circ}$F)              & --                                    \\ [.1cm]
\multirow{4}{*}{Reaction: Wood~\cite{SFPE:Tewarson}}       &  Formula: $\C\H_{1.7}\O_{0.74}\N_{0.002}$     & \multirow{4}{*}{Section~\ref{fire}}   \\
                                                           &  CO Yield: 0.004~$\mathrm{kg}/\mathrm{kg}$    &                                       \\
                                                           &  Soot Yield: 0.015~$\mathrm{kg}/\mathrm{kg}$  &                                       \\
                                                           &  Heat of Combustion: 16,400~kJ/kg             &                                       \\ [.25cm]
\multirow{2}{*}{Peak HRR}                                  &  Basement Fire: 6.75~MW                       & \multirow{2}{*}{Section~\ref{fire}}   \\ 
                                                           &  Bathroom Fire: 1.8~MW                        &                                       \\ [.25cm]                     
\multirow{3}{*}{Material: Wood~\cite{Incropera:1} }        &  $k$: 0.12~\si{W/(m.K)})                      & \multirow{3}{*}{Section~\ref{matl}}   \\
                                                           &  $\rho$: 510~kg/m$^3$                         &                                       \\
                                                           &  $c_{p}$: 1.38~\si{kJ/(kg.K)}                 &                                       \\ [.25cm]
\multirow{3}{*}{Material: Gypsum Board~\cite{WAKILI2007}}  &  $k$: 0.28~\si{W/(m.K)}                       & \multirow{3}{*}{Section~\ref{matl}}   \\ 
                                                           &  $\rho$: 810~kg/m$^3$                         &                                       \\ 
                                                           &  $c_{p}$: 1.0~\si{kJ/(kg.K)}                  &                                       \\
Wind 													   &  8.9~m/s (20~mph)							   & Section~\ref{Vents}                   \\
\bottomrule[1.25pt]
\end{tabular}\par
\footnotesize
$^{*}$Initial interior temperatures were assumed to be 20~$^{\circ}$C; exterior temperatures were 10~$^{\circ}$C~\cite{PGCounty2013}.
\normalsize
\end{table}

@@

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/PG_Total_9MW_HRR}
\caption[Interior prescribed and calculated interior HRRs vs. time from the simulation.]
{Comparison of prescribed and calculated interior HRRs from the FDS simulation. The vertical lines indicate the times when ventilation occurred.}
\label{fig:PG_Total_HRR}
\end{figure}

@@

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/PG_9MW_HRR}
\caption[Prescribed and total calculated HRRs vs. time from the simulation.]
{Comparison of total prescribed and calculated HRRs within the entire computational domain from the FDS simulation. The vertical lines indicate the times at which ventilation occurred.}
\label{fig:PG_9MW_HRR}
\end{figure}

@@

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/PG_Basement_9MW_HRR}
\caption[Prescribed, calculated domain, and calculated basement HRRs vs. time from the simulation.]
{Comparison of prescribed, total calculated domain HRR and basement calculated HRRs from the simulation. The vertical lines indicate the times when ventilation changed.}
\label{fig:PG_Basement_9MW_HRR}
\end{figure}

@@

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/PG_First_9MW_HRR}
\caption[First floor calculated HRR vs. time from the simulation.]
{First floor calculated HRR from the simulation. The vertical lines indicate the times when ventilation changed.}
\label{fig:PG_First_9MW_HRR}
\end{figure}

@@

\begin{figure}[!ht]
\includegraphics[trim = 1in 1in 1in 1in, clip=true, width=.8\textwidth]{../Figures/pressure_slice_99s}
\input{../Scripts/colorbar_metric_pres} \\
\vskip10pt
\includegraphics[trim = 1in 1in 1in 1in, clip=true, width=.8\textwidth]{../Figures/pressure_slice_101s}
\input{../Scripts/colorbar_metric_pres}
\caption[Calculated pressure on first floor 1~s before and after front door opens.]
{Calculated pressure contours on the first floor of the structure 1~s prior to (top) and 1~s after (bottom) firefighter entry through the front door at 0.91~m (3~ft) above the floor.}
\label{fig:pressure_slices}
\end{figure}

@@

\begin{figure}[!ht]
\includegraphics[trim = 0in 1in 0in 1in, clip=true, width=.65\textwidth]{../Figures/velocity_slice_99s}
\input{../Scripts/colorbar_metric_velo}
\input{../Scripts/colorbar_std_velo} \\
\includegraphics[trim = 0in 1in 0in 1in, clip=true, width=.65\textwidth]{../Figures/velocity_slice_101s}
\input{../Scripts/colorbar_metric_velo}
\input{../Scripts/colorbar_std_velo}
\caption[Calculated velocity on the first floor 1~s before and after front door opens]
{Calculated velocity contours on the first floor of the structure 1~s prior to (top) and 1~s after (bottom) firefighter entry through front door 0.91~m (3~ft) above the floor. Arrows indicate direction.}
\label{fig:velocity_slices}
\end{figure}

@@

\begin{figure}[!ht]
\includegraphics[trim = 1in 1in 1in 1in, clip=true, width=.65\textwidth]{../Figures/temperature_slice_99s}
\input{../Scripts/colorbar_metric_temp}
\input{../Scripts/colorbar_std_temp}
\caption[Calculated temperature on the first floor 1~s before front door opened.]
{Calculated temperature contours on the first floor of the structure 1~s prior to firefighter entry through front door 0.91~m (3~ft) above the floor.}
\label{fig:temperature_slice_99s}
\end{figure}

@@

\begin{figure}[!ht]
\includegraphics[trim = 1in 1in 1in 1in, clip=true, width=.65\textwidth]{../Figures/temperature_slice_101s}
\input{../Scripts/colorbar_metric_temp}
\input{../Scripts/colorbar_std_temp}
\caption[Calculated temperature on the first floor 1~s after front door opens.]
{Calculated temperature contours on the first floor of the structure 1~s after firefighter entry through front door 0.91~m (3~ft) above the floor.}
\label{fig:temperature_slice_101s}
\end{figure}

\begin{figure}[!ht]
\includegraphics[trim = 1in 1in 1in 1in, clip=true, width=.65\textwidth]{../Figures/temperature_slice_105s}
\input{../Scripts/colorbar_metric_temp}
\input{../Scripts/colorbar_std_temp}
\caption[Calculated temperature on first floor 5~s after front door opens]
{Calculated temperature contours on first floor of the structure 5~s after firefighter entry through front door 0.91~m (3~ft) above the floor.}
\label{fig:temperature_slice_105s}
\end{figure}

@@

\section{Assessing the Hazard}
\label{assessing_hazard}
A person is susceptible to second-degree burn injuries if skin temperatures exceed 55~$^{\circ}$C (130~$^{\circ}$F)~\cite{contactburn}. Although firefighters wear protective gear, gear only offers a finite amount of protection. The polycarbonate material in the facepiece of a self-contained breathing apparatus begins to soften when the material temperature reaches approximately 140~$^{\circ}$C (284~$^{\circ}$F)~\cite{mensch2011emergency}. Structural fire fighting coats and pants are tested to withstand temperatures up to 260~$^{\circ}$C (500~$^{\circ}$F) or 10~kW/m$^2$~\cite{nfpa2013standard}. Exposure to elevated temperatures can result in a significant amount of heat transferred to the firefighter, putting him or her at risk.

@@

\begin{figure}[!ht]
\begin{subfigure}{0.65\textwidth}
\includegraphics[trim = 0in 0in 0in 0in, clip=true, width=\textwidth]{../Figures/side_view_99s} \\
\vskip4pt
\includegraphics[trim = 0in 0in 0in 0in, clip=true, width=\textwidth]{../Figures/stair_view_99s} 
\end{subfigure}
\begin{subfigure}{0.35\textwidth}
\input{../Scripts/colorbar_metric_temp}
\input{../Scripts/colorbar_std_temp} \\
\end{subfigure}
\caption {Interior temperature contours 1~s before the front door was opened. The top snapshot shows temperatures on the first floor and basement on a plane that runs through the center of the front door. The bottom snapshot shows temperatures on the first floor and basement on the plane that runs through the center of the interior stairwell doorway.}
\label{fig:int_temp_99s}
\end{figure}

\begin{figure}[!ht]
\begin{subfigure}{0.65\textwidth}
\includegraphics[trim = 0in 0in 0in 0in, clip=true, width=\textwidth]{../Figures/side_view_101s} \\
\vskip4pt
\includegraphics[trim = 0in 0in 0in 0in, clip=true, width=\textwidth]{../Figures/stair_view_101s} 
\end{subfigure}
\begin{subfigure}{0.35\textwidth}
\input{../Scripts/colorbar_metric_temp}
\input{../Scripts/colorbar_std_temp} \\
\end{subfigure}
\caption{Interior temperature contours 1~s after the front door was opened. The top snapshot shows temperatures on the first floor and basement on a plane that runs through the center of the front door. The bottom snapshot shows temperatures on the first floor and basement on the plane that runs through the center of the interior stairwell doorway.}
\label{fig:int_temp_101s}
\end{figure}

\begin{figure}[!ht]
\begin{subfigure}{0.65\textwidth}
\includegraphics[trim = 0in 0in 0in 0in, clip=true, width=\textwidth]{../Figures/side_view_142s} \\
\vskip4pt
\includegraphics[trim = 0in 0in 0in 0in, clip=true, width=\textwidth]{../Figures/stair_view_142s} 
\end{subfigure}
\begin{subfigure}{0.35\textwidth}
\input{../Scripts/colorbar_metric_temp}
\input{../Scripts/colorbar_std_temp} \\
\end{subfigure}
\caption{Interior temperature contours as firefighters made entry to the structure. The top snapshot shows temperatures on the first floor and basement on a plane that runs through the center of the front door. The bottom snapshot shows temperatures on the first floor and basement on the plane that runs through the center of the interior stairwell doorway.}
\label{fig:int_temp_142s}
\end{figure}

@@

The interior stairwell acted as a chimney for hot gases in the basement to flow towards regions of lower pressure and vent openings located on the front side of the structure. Firefighters should avoid placing themselves within a flow path where elevated temperatures and flow velocities can present hazardous conditions and increase the rate of heat transfer to firefighting gear via convection. A 360$^\circ$ scene size-up by arriving firefighters can help determine the location of the fire and identify potential flow paths within a structure. Door control can also be used to avoid creating inlet and outlet vents that could result in the establishment of a flow path.

First suppression efforts should take place at the same level as the first with the wind at the back of firefighters. Ongoing research by NIST, Underwriters Laboratories (UL), and others has demonstrated that applying water from the exterior into the fire area of a structure (prior to the start of interior operations) can significantly improve the safety of firefighters by reducing the thermal hazard from the fire and reducing the potential for developing high velocity hot gas flows within the structure~\cite{madrzykowski2009fire, kerber2009fire}.

@@

Based on a review of these incidents, it is clear that fires with rapidly developing or changing flow paths are a significant hazard to the fire service. The development of (or changes to) a flow path could be caused by the failure of a component of the structure, such as a door, window, or portion of a ceiling, wall or floor. Environmental conditions such as wind can generate hazardous thermal conditions within a flow path. Uncoordinated ventilation procedures can also be the cause of increased thermal hazards within a flow path.

>>>>

\large
\flushright{\bf National Institute of Standards and Technology Technical Note XXXX \\
Natl.~Inst.~Stand.~Technol.~Tech.~Note~XXXX, \pageref{LastPage} pages (February 2015) \\
CODEN: NTNOEF}

@@

\begin{table}[!h]
\centering
\captionof{table}[Abridged approximate fire event timeline.]{Abridged Prince George's Fire/EMS approximate fire event timeline~\cite{PGCounty2013}.}\label{tab:fire_info}
\begin{tabular}{cl}
\toprule[1.5pt]
Incident Time              &  Fire Behavior / Fireground Operation  \\
{[hh:mm:ss]}               &                                        \\
\midrule
21:08:26   & First 911 call is made. \\[.25cm]
\multirow{3}{*}{21:11:03}  &   \multirow{3}{*}{\parbox{10cm} {Dispatch for a ``box'' alarm assignment- four engine companies, two truck companies, one rescue squad, and a commanding officer.}}    \\
                           &         \\
                           &         \\[.25cm]
21:12:55  & First unit (Engine 807B) arrived on-scene.  \\[.25cm]
21:13:44  & Volunteer Chief 809A establishes incident command (IC).   \\[.25cm]
21:13:56  & Truck 809 arrived on-scene. \\[.25cm]
21:14:00  & Front door opened. \\[.25cm]
21:14:42  & Truck 809 personnel enter structure. \\[.25cm]
21:15:47  & Basement door opened and water applied to fire. Front door blown closed. \\[.25cm]
21:15:51  & Lower pane of front egress window broken. \\[.25cm]
21:16:01  & Top pane of front egress window broken. \\[.25cm]
21:16:16  & Truck 809 officer bailout through front window. \\[.25cm]
21:16:44  & Front door opened. \\[.25cm]
21:17:15  & Truck 809 forcible entry firefighter rescued. \\[.25cm]
\bottomrule[1.25pt]
\end{tabular}\par
\end{table}

@@

\begin{figure}[!ht]
\includegraphics[width=0.75\textwidth]{../Figures/Fire_HRR}
\caption[Prescribed HRR vs. time for the simulation.]
{Prescribed HRR vs. time for the simulation. The text labels indicate the prescribed steady peak HRR and the decay phase associated with firefighter water application. Time zero represnts 21:21:20, see Table~\ref{tab:combined_timeline}}.
\label{fig:hrr}
\end{figure}

@@

In addition to the state of openings within the structure, outlined in Table~\ref{tab:vents}, an 8.9~m/s (20~mph) wind velocity boundary condition was assumed. The flow direction of the wind was asasumed from the rear toward the front of the structure, perpendicular to the rear of the structure.  Based on local weather conditions, mean wind velocities ranged between 4.5~m/s (10~mph) and 8.9~m/s (20~mph) with gusts up to 13.4~m/s (30~mph) with a direction that angled toward the back left of the structure~\cite{PGCounty2013}. The following lines define the wind boundary condition in the FDS input file:

\begin{lstlisting}
&SURF ID='wind', VEL=-8.94, RAMP_V='WIND RAMP',COLOR='INVISIBLE' /
&VENT MB = 'YMAX', SURF_ID ='wind' /
&RAMP ID='WIND RAMP', T=0,F=0.25 /
&RAMP ID='WIND RAMP', T=10,F=0.50 /
&RAMP ID='WIND RAMP', T=20,F=0.75 /
&RAMP ID='WIND RAMP', T=30,F=1.00 /
\end{lstlisting}

Note that there is a 30~s ramp time to reach peak wind speed. This was inlcuded to prevent an abrupt change at the FDS domain boundary. Since the front door does not open until 100~s into the simulation there was sufficient time to for the airflow to stabilize.

\begin{table}
\centering
\captionof{table}{Timeline of ventilation changes in the simulation.}\label{tab:vents}
\begin{tabular}{lllll}
\toprule[1.5pt]
Baseline     &   Ventilation   &  Type       & State of       &  Side of  \\
Simulation  &  Area             &                 & Opening     &  Structure  \\
{[s]}             &  {[m$^2$]}     &                 &           &               \\
\midrule
0               &  1.40             &  Window      &  Opened     &  Rear  \\[.25cm]
0               &  0.39             &  Window      &  Opened     &  Left  \\[.25cm]
100             &  1.84             &  Door        &  Opened     &  Front \\[.25cm]
207             &  1.84             &  Door        &  Closed     &  Front \\[.25cm]
207             &  1.84             &  Door        &  Opened     &  Rear  \\[.25cm]
211             &  0.26             &  Window      &  Opened     &  Front \\[.25cm]
221             &  0.26             &  Window      &  Opened     &  Front \\[.25cm]
267             &  1.84             &  Door        &  Opened     &  Front \\[.25cm]

@@

\begin{table}[!ht]
\centering
\captionof{table}{Relevant fire model input parameters.}\label{tab:mod_param}
\begin{tabular}{lll}
\toprule[1.5pt]
Parameter                                                  & Description                                   & Discussion                            \\
\midrule
Simulation Time                                            &  5~min                                        & --                                    \\ [.25cm]
\multirow{2}{*}{Grid Cell Size}                            &  10~cm around house                           & \multirow{2}{*}{Section~\ref{mesh}}   \\ 
                                                           &  20~cm for wind                               &                                       \\ [.25cm]
Ambient Temperature$^*$                                    &  20~$^{\circ}$C (68~$^{\circ}$F)              & --                                    \\ [.1cm]
\multirow{4}{*}{Reaction: Wood~\cite{SFPE:Tewarson}}       &  Formula: $\C\H_{1.7}\O_{0.74}\N_{0.002}$     & \multirow{4}{*}{Section~\ref{fire}}   \\
                                                           &  CO Yield: 0.004~$\mathrm{kg}/\mathrm{kg}$    &                                       \\
                                                           &  Soot Yield: 0.015~$\mathrm{kg}/\mathrm{kg}$  &                                       \\
                                                           &  Heat of Combustion: 16,400~kJ/kg             &                                       \\ [.25cm]
\multirow{2}{*}{Peak HRR}                                  &  Basement Fire: 6.75~MW                       & \multirow{2}{*}{Section~\ref{fire}}   \\ 
                                                           &  Bathroom Fire: 1.8~MW                        &                                       \\ [.25cm]                     
\multirow{3}{*}{Material: Wood~\cite{Incropera:1} }        &  $k$: 0.12~\si{W/(m.K)})                      & \multirow{3}{*}{Section~\ref{matl}}   \\
                                                           &  $\rho$: 510~kg/m$^3$                         &                                       \\
                                                           &  $c_{p}$: 1.38~\si{kJ/(kg.K)}                 &                                       \\ [.25cm]
\multirow{3}{*}{Material: Gypsum Board~\cite{WAKILI2007}}  &  $k$: 0.28~\si{W/(m.K)}                       & \multirow{3}{*}{Section~\ref{matl}}   \\ 
                                                           &  $\rho$: 810~kg/m$^3$                         &                                       \\ 
                                                           &  $c_{p}$: 1.0~\si{kJ/(kg.K)}                  &                                       \\ [.25cm]
Wind 													   &  8.9~m/s (20~mph)							   & Section~\ref{Vents}                   \\
\bottomrule[1.25pt]
\end{tabular}\par
\footnotesize
$^{*}$Initial interior temperatures were assumed to be 20~$^{\circ}$C; exterior temperatures were 10~$^{\circ}$C~\cite{PGCounty2013}.
\normalsize
\end{table}

@@

\begin{figure}[!ht]
\includegraphics[width=\textwidth]{../Figures/PG_Total_9MW_HRR}
\caption[Interior prescribed and calculated interior HRRs vs. time from the simulation.]
{Comparison of prescribed and calculated interior HRRs from the FDS simulation. The vertical lines indicate the times when ventilation occurred.}
\label{fig:PG_Total_HRR}
\end{figure}

@@

\begin{figure}[!ht]
\includegraphics[width=\textwidth]{../Figures/PG_9MW_HRR}
\caption[Prescribed and total calculated HRRs vs. time from the simulation.]
{Comparison of total prescribed and calculated HRRs within the entire computational domain from the FDS simulation. The vertical lines indicate the times at which ventilation occurred.}
\label{fig:PG_9MW_HRR}
\end{figure}

@@

\begin{figure}[!ht]
\includegraphics[width=\textwidth]{../Figures/PG_Basement_9MW_HRR}
\caption[Prescribed, calculated domain, and calculated basement HRRs vs. time from the simulation.]
{Comparison of prescribed, total calculated domain HRR and basement calculated HRRs from the simulation. The vertical lines indicate the times when ventilation changed.}
\label{fig:PG_Basement_9MW_HRR}
\end{figure}

@@

\begin{figure}[!ht]
\includegraphics[width=\textwidth]{../Figures/PG_First_9MW_HRR}
\caption[First floor calculated HRR vs. time from the simulation.]
{First floor calculated HRR from the simulation. The vertical lines indicate the times when ventilation changed.}
\label{fig:PG_First_9MW_HRR}
\end{figure}

@@

\begin{figure}[!ht]
\includegraphics[trim = 1in 1in 1in 1in, clip=true, width=.8\textwidth]{../Figures/pressure_slice_99s}
\input{../Scripts/colorbar_metric_pres} \\
\vskip10pt
\includegraphics[trim = 1in 1in 1in 1in, clip=true, width=.8\textwidth]{../Figures/pressure_slice_101s}
\input{../Scripts/colorbar_metric_pres}
\caption[Calculated pressure on first floor 1~s before and after the front door opened.]
{Calculated pressure contours on the first floor of the structure 1~s prior to (top) and 1~s after (bottom) firefighter entry through the front door at 0.91~m (3~ft) above the floor.}
\label{fig:pressure_slices}
\end{figure}

@@

\begin{figure}[!ht]
\includegraphics[trim = 0in 1in 0in 1in, clip=true, width=.65\textwidth]{../Figures/velocity_slice_99s}
\input{../Scripts/colorbar_metric_velo}
\input{../Scripts/colorbar_std_velo} \\
\includegraphics[trim = 0in 1in 0in 1in, clip=true, width=.65\textwidth]{../Figures/velocity_slice_101s}
\input{../Scripts/colorbar_metric_velo}
\input{../Scripts/colorbar_std_velo}
\caption[Calculated velocity on the first floor 1~s before and after the front door opened.]
{Calculated velocity contours on the first floor of the structure 1~s prior to (top) and 1~s after (bottom) firefighter entry through front door 0.91~m (3~ft) above the floor. Arrows indicate direction.}
\label{fig:velocity_slices}
\end{figure}

@@

\begin{figure}[!ht]
\includegraphics[trim = 1in 1in 1in 1in, clip=true, width=.65\textwidth]{../Figures/temperature_slice_99s}
\input{../Scripts/colorbar_metric_temp}
\input{../Scripts/colorbar_std_temp}
\caption[Calculated temperature on the first floor 1~s before the front door opened.]
{Calculated temperature contours on the first floor of the structure 1~s prior to firefighter entry through front door 0.91~m (3~ft) above the floor.}
\label{fig:temperature_slice_99s}
\end{figure}

@@

\begin{figure}[!ht]
\includegraphics[trim = 1in 1in 1in 1in, clip=true, width=.65\textwidth]{../Figures/temperature_slice_101s}
\input{../Scripts/colorbar_metric_temp}
\input{../Scripts/colorbar_std_temp}
\caption[Calculated temperature on the first floor 1~s after the front door opened.]
{Calculated temperature contours on the first floor of the structure 1~s after firefighter entry through front door 0.91~m (3~ft) above the floor.}
\label{fig:temperature_slice_101s}
\end{figure}

\begin{figure}[!ht]
\includegraphics[trim = 1in 1in 1in 1in, clip=true, width=.65\textwidth]{../Figures/temperature_slice_105s}
\input{../Scripts/colorbar_metric_temp}
\input{../Scripts/colorbar_std_temp}
\caption[Calculated temperature on first floor 5~s after the front door opened.]
{Calculated temperature contours on first floor of the structure 5~s after firefighter entry through front door 0.91~m (3~ft) above the floor.}
\label{fig:temperature_slice_105s}
\end{figure}

@@

\section{Assessing the Hazard}
\label{assessing_hazard}
A person is susceptible to second-degree burn injuries if skin temperatures exceed 55~$^{\circ}$C (130~$^{\circ}$F)~\cite{contactburn}. Although firefighters wear protective gear, gear only offers a finite amount of protection. The polycarbonate material in the facepiece of a self-contained breathing apparatus begins to soften when the material temperature reaches approximately 140~$^{\circ}$C (284~$^{\circ}$F)~\cite{mensch2011emergency}. Structural fire fighting coats and pants are tested to withstand temperatures up to 260~$^{\circ}$C (500~$^{\circ}$F) or 10~kW/m$^2$~\cite{nfpa2013standard} for less than 2.5~min. Exposure to elevated temperatures can result in a significant amount of heat transferred to the firefighter, putting him or her at risk.

@@

\begin{figure}[!ht]
\begin{subfigure}{0.65\textwidth}
\includegraphics[trim = 0in 0in 0in 0in, clip=true, width=\textwidth]{../Figures/side_view_99s} \\
\vskip4pt
\includegraphics[trim = 0in 0in 0in 0in, clip=true, width=\textwidth]{../Figures/stair_view_99s} 
\end{subfigure}
\begin{subfigure}{0.35\textwidth}
\input{../Scripts/colorbar_metric_temp}
\input{../Scripts/colorbar_std_temp} \\
\end{subfigure}
\caption {Interior temperature contours 1~s before the front door was opened (21:13:59). The top snapshot shows temperatures on the first floor and basement on a plane that runs through the center of the front door. The bottom snapshot shows temperatures on the first floor and basement on the plane that runs through the center of the interior stairwell doorway.}
\label{fig:int_temp_99s}
\end{figure}

\begin{figure}[!ht]
\begin{subfigure}{0.65\textwidth}
\includegraphics[trim = 0in 0in 0in 0in, clip=true, width=\textwidth]{../Figures/side_view_101s} \\
\vskip4pt
\includegraphics[trim = 0in 0in 0in 0in, clip=true, width=\textwidth]{../Figures/stair_view_101s} 
\end{subfigure}
\begin{subfigure}{0.35\textwidth}
\input{../Scripts/colorbar_metric_temp}
\input{../Scripts/colorbar_std_temp} \\
\end{subfigure}
\caption{Interior temperature contours 1~s after the front door was opened (21:14:01). The top snapshot shows temperatures on the first floor and basement on a plane that runs through the center of the front door. The bottom snapshot shows temperatures on the first floor and basement on the plane that runs through the center of the interior stairwell doorway.}
\label{fig:int_temp_101s}
\end{figure}

\begin{figure}[!ht]
\begin{subfigure}{0.65\textwidth}
\includegraphics[trim = 0in 0in 0in 0in, clip=true, width=\textwidth]{../Figures/side_view_142s} \\
\vskip4pt
\includegraphics[trim = 0in 0in 0in 0in, clip=true, width=\textwidth]{../Figures/stair_view_142s} 
\end{subfigure}
\begin{subfigure}{0.35\textwidth}
\input{../Scripts/colorbar_metric_temp}
\input{../Scripts/colorbar_std_temp} \\
\end{subfigure}
\caption{Interior temperature contours as firefighters made entry to the structure (21:14:42). The top snapshot shows temperatures on the first floor and basement on a plane that runs through the center of the front door. The bottom snapshot shows temperatures on the first floor and basement on the plane that runs through the center of the interior stairwell doorway.}
\label{fig:int_temp_142s}
\end{figure}

@@

The interior stairwell acted as a chimney for hot gases in the basement to flow towards regions of lower pressure and vent openings located on the front side of the structure. Firefighters should avoid placing themselves within a flow path where elevated temperatures and flow velocities can present hazardous conditions and increase the rate of heat transfer to the firefighter via convection. A 360$^\circ$ scene size-up by arriving firefighters can help determine the location of the fire and identify potential flow paths within a structure. Door control can also be used to avoid creating inlet and outlet vents that could result in the establishment of a flow path.

First suppression efforts should take place at the same level as the fire (i.e. fight basement fires from the basement level) with the wind at the back of firefighters. Ongoing research by NIST, Underwriters Laboratories (UL), and others has demonstrated that applying water from the exterior into the fire area of a structure (prior to the start of interior operations) can significantly improve the safety of firefighters by reducing the thermal hazard from the fire and reducing the potential for developing high velocity hot gas flows within the structure~\cite{madrzykowski2009fire, kerber2009fire}.

@@

Based on a review of these incidents, it is clear that fires with rapidly developing or changing ventilation may lead to flow paths that are a significant hazard to the fire service during a response. The development of (or changes to) a flow path could be caused by the failure of a component of the structure, such as a door, window, or portion of a ceiling, wall or floor. Environmental conditions such as wind can generate hazardous thermal conditions within a flow path. Uncoordinated ventilation procedures can also be the cause of increased thermal hazards within a flow path.

\begin{table}[!ht]
\caption[Flow path related LODD/LODI incidents.]
{Flow path related LODD/LODI incidents.}
\begin{tabular}{lll}
\toprule
NIOSH Report No. & No. of LODDs/LODIs & Flow Path Details                                              \\
\midrule
99-F01   \cite{NIOSH:Pettit}        &  3 LODDs            &  From apartment into hallway on 10th       \\
                                    &                     &  floor of high-rise apartment building     \\
99-F21   \cite{NIOSH:Washenitz}     &  2 LODDs            &  Basement to 1st floor                     \\
                                    &  2 LODIs            &                                            \\
F2000-04 \cite{NIOSH:Mezzanotte}    &  3 LODDs            &  1st floor to 2nd floor                    \\
                                    &  3 civilian deaths  &                                            \\
F2000-16 \cite{NIOSH:McFall}        &  1 LODD             &  2nd floor hallway through                 \\
                                    &  1 LODI             &  2nd floor apartment                       \\
                                    &  1 civilian death   &                                            \\
F2000-23 \cite{NIOSH:McFall2}       &  1 LODD             &  From ground level to 1st floor then to    \\
                                    &  2 LODIs            &  2nd floor, flow exited through ceiling    \\
F2000-43 \cite{NIOSH:McFall3}       &  1 serious LODI     &  1st floor to 2nd floor                    \\
                                    &  2 other LODIs      &                                            \\
F2004-02 \cite{NIOSH:Berardinelli}  &  1 LODD             &  1st floor to basement                     \\
F2005-02 \cite{NIOSH:Koedam}        &  1 LODD             &  Rear to front of the building             \\
                                    &  4 LODIs            &                                            \\
F2005-04 \cite{NIOSH:McFall4}       &  1 LODD             &  Basement to 1st floor                     \\
                                    &  9 LODIs            &                                            \\
F2007-09 \cite{NIOSH:Tarley}        &  1 LODD             &  3 story training burn - flow through      \\
                                    &  2 LODIs            &  all levels                                \\
F2007-35 \cite{NIOSH:Braddee}       &  4 LODIs            &  1st floor to 2nd floor                    \\
F2009-11 \cite{NIOSH:Merinar}       &  2 LODDs            &  Rear to front of the building             \\
F2011-13 \cite{NIOSH:Bowyer2}       &  2 LODDs            &  Lower level up stairs and through         \\
                                    &                     &  entry door and garage                     \\
F2011-31 \cite{NIOSH:Loflin}        &  1 LODD             &  Fire extended from lower level apartment  \\
F2012-28 \cite{NIOSH:Bowyer}        &  1 LODD             &  Attic fire extended into closed           \\
                                    &  1 LODI             &  porch and then into 2nd floor             \\
\bottomrule
\end{tabular}
\label{tab:LODD}
\end{table}
